---
title: "Activity 10 - The Chaos Game"
author: "Dr. VanKrevelen -- STS3470 (Spring 2023)"
date: 'Updated: `r Sys.Date()`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:** Use this .RMD script to record your code / output / answers.

Remember that with a .RMD file, you can run individual code chunks by clicking the green triangle button in the top right of the chunk or by putting your cursor on a line and hitting Cmd+Return (Mac) or Ctrl+Return (PC) like you would in a .R script. When you use the Knit button, it will run all the code in order in a fresh R session and produce your knitted file.

Have questions? Find a classmate or Dr. V to ask!

Below is code to set up the Chaos Game (from the notes):

```{r}
# Set up Graph

plot(0:2, 0:2, type = "n")
x <- c(0, 1, 2)
y <- c(0, 2, 0)
points(x, y, pch = 19)
labels <- c("A", "B", "C")
text(x + 0.05, y, labels)
segments(0, 0, 2, 0)
segments(0, 0, 1, 2)
segments(1, 2, 2, 0)
points(1, 1, pch = 19)


# Determine starting point and vertices to choose from

myX <- 1
myY <- 1
myvertices <- c("A", "B", "C")


# Run Chaos Game

for(i in 1:1000) {
  mycorner <- sample(myvertices, 1)
  
  if(mycorner == "A") {
    myX <- (myX + 0) / 2
    myY <- (myY + 0) / 2
  } else if(mycorner == "B") {
    myX <- (myX + 1) / 2
    myY <- (myY + 2) / 2
  } else {
    myX <- (myX + 2) / 2
    myY <- (myY + 0) / 2
  }
  points(myX, myY, pch = 19)
}
```

Use the code and output to answer the following questions. If you ever aren't sure, try doing what it asks and see if you can figure out what's changing and why. You can also change how many iterations you have to something smaller than 10000 if you want to see how the code works over a smaller number of iterations.

<br>
<br>


1) Why is it important that I put my initial `myX <- 1` and `myY <- 1` outside of the for loop? In other words, what would happen if I put it inside the loop where the code for `mycorner` is?

**Answer:** If that code was inside of the loop, I would reset my point to (1, 1) at the start of each iteration, and my points would never move throughout the graph.

<br>
<br>


2) Would it matter if I moved the code `myvertices <- c("A", "B", "C")` inside my for loop? What would change about how the code runs?

**Answer:** It would not change any results. However, it would cause this code to run during each iteration of the loop, which is unneccessary. I only need to create the object once, so it makes more sense to include it before the loop.

<br>
<br>


3) What would happen if the `mycorner <- sample(myvertices, 1)` code were moved before the for loop (like `myvertices`)?

**Answer:** I would only choose a random corner once (instead of each iteration of the loop), so my points would just keep getting closer and closer to the one chosen corner.

<br>
<br>


4) What would happen if the `points(myX, myY, pch = 19)` code were placed outside the for loop at the end instead of inside of it?

**Answer:** I would only add one point to my graph (at the very end of the loop) instead of adding new points each time I chose a new randomly selected corner.

<br>
<br>


5) What is happening to the values for `myX` and `myY` during each iteration of the loop (i.e. when i = 1, then 2, then 3, etc.)?

**Answer:** Each time i takes on a new value I am updating `myX` and `myY` to move them halfway from their current value to the values for the new corner I randomly selected.

<br>
<br>


**Bonus:** What would happen if you updated the code to do one of the following?

- Make it so that you can't go toward the same corner twice in a row

**Answer:** The pattern changes to really limit where the points are able to appear.

```{r}
library(dplyr) # to use case_when()

# Set up Graph

plot(0:2, 0:2, type = "n")
x <- c(0, 1, 2)
y <- c(0, 2, 0)
points(x, y, pch = 19)
labels <- c("A", "B", "C")
text(x + 0.05, y, labels)
segments(0, 0, 2, 0)
segments(0, 0, 1, 2)
segments(1, 2, 2, 0)
points(1, 1, pch = 19)


# Determine starting point and vertices to choose from

myX <- 1
myY <- 1
myvertices <- c("A", "B", "C")


# Run Chaos Game

for(i in 1:1000) {
  mycorner <- case_when(mycorner == "A" ~ sample(c("B", "C"), 1),
                        mycorner == "B" ~ sample(c("A", "C"), 1),
                        mycorner == "C" ~ sample(c("A", "B"), 1))

  if(mycorner == "A") {
    myX <- (myX + 0) / 2
    myY <- (myY + 0) / 2
  } else if(mycorner == "B") {
    myX <- (myX + 1) / 2
    myY <- (myY + 2) / 2
  } else {
    myX <- (myX + 2) / 2
    myY <- (myY + 0) / 2
  }
  points(myX, myY, pch = 19)
}
```

- Start with a square instead of a triangle

**Answer:** It looks mostly like random scatter.

```{r}
# Set up Graph

plot(0:1, 0:1, type = "n")
x <- c(0, 0, 1, 1)
y <- c(0, 1, 1, 0)
points(x, y, pch = 19)
labels <- c("A", "B", "C", "D")
text(x + 0.02, y + 0.02, labels)
segments(0, 0, 0, 1)
segments(0, 1, 1, 1)
segments(1, 1, 1, 0)
segments(1, 0, 0, 0)
points(0.5, 0.5, pch = 19)


# Determine starting point and vertices to choose from

myX <- 0.5
myY <- 0.5
myvertices <- c("A", "B", "C", "D")


# Run Chaos Game

for(i in 1:1000) {
  mycorner <- sample(myvertices, 1)
  
  if(mycorner == "A") {
    myX <- (myX + 0) / 2
    myY <- (myY + 0) / 2
  } else if(mycorner == "B") {
    myX <- (myX + 0) / 2
    myY <- (myY + 1) / 2
  } else if (mycorner == "C") {
    myX <- (myX + 1) / 2
    myY <- (myY + 1) / 2
  } else {
    myX <- (myX + 1) / 2
    myY <- (myY + 0) / 2
  }
  
  points(myX, myY, pch = 19)
}
```

- Some other variation you can think of. You could try a square where you can't go to the same corner twice in a row.