---
title: "Activity 10A - Conditional Computing"
author: "Dr. VanKrevelen -- STS3470 (Spring 2023)"
date: 'Updated: `r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:** Use this .RMD script to record your code / output / answers.

Remember that with a .RMD file, you can run individual code chunks by clicking the green triangle button in the top right of the chunk or by putting your cursor on a line and hitting Cmd+Return (Mac) or Ctrl+Return (PC) like you would in a .R script. When you use the Knit button, it will run all the code in order in a fresh R session and produce your knitted file.

Have questions? Find a classmate or Dr. V to ask!



1. Below is code that calculates a p-value for a t-test and then uses if statements to update an output. Run the code and answer the questions associated with it.

```{r}
set.seed(9876)
pval <- t.test(rnorm(30, 0, 5), mu = -2)$p.value
status <- NA

if(pval < 0.05){
  status <-  "Reject null"
}

status
```

a. What is the true value of $\mu$?

**Answer:** 0 (sampling from a normal distribution with mean 0)

<br>


b. What hypotheses are being tested?

**Answer:** $H_0: \mu = -2$ vs. $H_a: \mu \ne -2$

<br>


c. What is our p-value?

**Answer:** 0.015

<br>


d. What value does status have at the end?

**Answer:** Reject null

<br>


2. Copy and paste the code from #1 below but use a seed of 1234.

```{r}
set.seed(1234)
pval <- t.test(rnorm(30, 0, 5), mu = -2)$p.value
status <- NA

if(pval < 0.05){
  status <-  "Reject null"
}

status
```

a. What is our p-value now?

**Answer:** 0.535

<br>


b. What value does status have now?

**Answer:** NA (it didn't change when I ran my if statement)

<br>


3. Rewrite this code using `ifelse()` and assigning a value of "Fail to reject null" if our p-value is > 0.05. Test it to verify you can get both answers.

```{r}
set.seed(1234)
pval <- t.test(rnorm(30, 0, 5), mu = -2)$p.value
status <- NA
status <- ifelse(pval > 0.05, "Fail to reject null", "Reject null")
status
```

<br>


4. Try using `case_when()` to make status say "Reject null" if 0 < pval < 0.05, "Fail to reject null" if 0.05 < pval < 1, and "ERROR" if pval < 0 or pval > 1.

```{r, message = FALSE}
library(dplyr) # needed for case_when()

pval <- 0.34

case_when(pval < 0 ~ "ERROR",
          pval < 0.05 ~ "Reject null",
          pval <= 1 ~ "Fail to reject null",
          TRUE ~ "ERROR")
```

<br>


5. Many courses use scores in the 90s as an A, 80s as a B, 70s as a C, 60s as a D, and below 60 as an F. Complete the code below by filling in the blanks with appropriate symbols. 

```{r, eval = TRUE}
average <- 77
if(average >= 90 & average <= 100) { grade <- "A" }
if(average >= 80 & average < 90) { grade <- "B" }
if(average >= 70 & average < 80) { grade <- "C" }
if(average >= 60 & average < 70) { grade <- "D" }
if(average < 60) { grade <- "F" }
grade
```

a. Change `eval = FALSE` to `TRUE` when you are done and verify grade is assigned "C" for an average of 77.

<br>


b. Change average to 87 and re-run the code. Then change average to 70. Do you get the appropriate output both times?

**Answer:** Yes

<br>


c. What would happen if a gradebook incorrectly had an average of 103 for someone? Try it out.

**Answer:** It will report whatever grade was last calculated because 103 doesn't fall into any of the logical statements we created inside `if()`. In this case, it says "C" because I last checked an average of 70.

<br>


d. What if a typo listed someone's grade as -95 instead of 95?

**Answer:** The same thing will happen. It will report the previous grade.

<br>


e. What could you do to improve the code to avoid the problems in part c/d? Try it out and see if it works.

```{r}
average <- 103
if(average >= 90 & average <= 100) { grade <- "A" }
if(average >= 80 & average < 90) { grade <- "B" }
if(average >= 70 & average < 80) { grade <- "C" }
if(average >= 60 & average < 70) { grade <- "D" }
if(average < 60) { grade <- "F" }
if(average > 100 | average < 0) { 
  grade <- "Grade is negative or over 100. Is there a typo?" }
grade
```

**Answer:** We could add another if statement that produces a warning message.

<br>


f. What happens if you have a vector of grades like the one below?

```{r, eval = FALSE}
average <- c(87, 100, 92, 74)
if(average >= 90 & average <= 100) { grade <- "A" }
if(average >= 80 & average < 90) { grade <- "B" }
if(average >= 70 & average < 80) { grade <- "C" }
if(average >= 60 & average < 70) { grade <- "D" }
if(average < 60) { grade <- "F" }
if(average > 100 | average < 0) { 
  grade <- "Grade is negative or over 100. Is there a typo?" }
grade
```

**Answer:** This produces an error saying my condition has length > 1.

<br>


g. (Bonus) Try recreating this code using `case_when()`. Does it work with vectors?

```{r}
average <- c(87, 100, 92, 74, -84, 62, 45, 192)
grade <- case_when(average > 100 ~ "Error. Grade was above 100",
                   average >= 90 ~ "A",
                   average >= 80 ~ "B",
                   average >= 70 ~ "C",
                   average >= 60 ~ "D",
                   average >= 0 ~ "F",
                   TRUE ~ "Error. Grade was negative.")
grade
```

**Answer:** Yes, `case_when()` does work with vectors!

<br>


