---
title: 'Activity 11 - Writing R Functions'
author: "Dr. VanKrevelen -- STS3470 (Spring 2023)"
date: 'Updated: `r Sys.Date()`'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:** Use this .RMD script to record your code / output / answers.

Remember that with a .RMD file, you can run individual code chunks by clicking the green triangle button in the top right of the chunk or by putting your cursor on a line and hitting Cmd+Return (Mac) or Ctrl+Return (PC) like you would in a .R script. When you use the Knit button, it will run all the code in order in a fresh R session and produce your knitted file.

Have questions? Find a classmate or Dr. V to ask!

# Part A - Practice Creating Functions

The code below can be used to explore Type I Errors (aka false positives). It uses a pre-specified number or samples, sample size, null value, and seed.

```{r}
set.seed(2320)
mydata <- matrix(rnorm(1000 * 30), nrow = 1000, ncol = 30)
mypvals <- NULL
for(i in 1:1000) {
  out <- t.test(mydata[i, ], alternative = "two.sided", mu = 0)
  mypvals[i] <- out$p.value
}

sum(mypvals <= 0.05) / 1000
```

Let's write a function that let's a user choose different values for the simulation.

1. Create a function called `rejectH0()` that takes the previous code and has the following arguments:

    - samples (default of 1000)
    - samplesize (default of 30)
    - munull (default of 0)
    - seed (default of 2320)

Try running the function with no arguments specified and verify that you get the same value (0.051) that we got above.

```{r}

```

<br>


2. Use your function to draw 5,000 samples of size 20 from a standard normal distribution and test $H_0: \mu = 0$ vs. $H_a: \mu \ne 0$. What proportion do you get? Is this surprising or what we would expect? Justify your answer.

```{r}

```

**Answer:**

<br>


3. Now draw the same 5,000 samples of size 20 but test $H_0: \mu = 1$ vs. $H_a: \mu \ne 1$. What proportion do we get this time? Explain why this answer makes sense when compared with your previous answer.

```{r}

```

**Answer:**

<br>


4. What is the name for the proportion from the previous question? In other words, we would say the test has a __________________ of 0.9884.

**Answer:**

<br>


5. Edit our function so that it includes a fifth argument called `mutrue` with a default value of 0 that can be used to change the mean of the population from which our sample was drawn.

```{r}

```

<br>


6. Use the function to draw 5,000 samples of size 10 from a normal distribution with mean 0.5 (and standard deviation 1 still) and to test $H_0: \mu = 0$ vs. $H_a: \mu \ne 0$. How does this answer compare to our previous answers and why do we get a number like this?

```{r}

```

**Answer:**

<br>


7. What would you expect to happen as you increase the sample size but keep the other elements of the function the same? Try testing this for a couple of new sample sizes.

```{r}

```

**Answer:**

<br>


8. (Bonus) Could you write a for loop that uses your function to test samples of size 10, 20, 30, 40, and 50?

```{r}

```

<br>

9. Write a function called `my_sum()` that contains a single integer argument. The function should compute the sum of all integers from 1 up to your argument. For exmaple `my_sum(3)` would calculate 1 + 2 + 3 and would output 6. Verify that your function works by testing a couple of examples. `my_sum(10)` should give you a value of 55 and `my_sum(100)` should give you a value of 5050.

```{r}

```

---

# Part B - More practice with functions

Many board games involve rolling one or more dice and recording a result like the sum of the rolls, the maximum of the rolls, or the minimum of the rolls. We can simulate rolling `k` dice with `n` sides by using the sample function as `sample(1:n, k, replace = TRUE)`.

10. Write a function called `roll_dice()` with three arguments: dice, sides, type. The arguments should have respective default values of 2, 6, and "sum". The dice and sides arguments will be used in the `sample()` function example above. You should use conditional computing to output either the sum of the dice, the maximum of the dice, or the minimum of the dice (based on whether `type` is "sum", "min", or "max").

```{r}

```

<br>


11. Games like Monopoly involve rolling two six-sided dice and using the sum to decide how far you move. Write a for loop that rolls two six-sided dice 1,000 times and records the sum of the rolls each time. Then create a graph displaying how often each sum occurs. What do you notice?

```{r}

```

**Answer:**

<br>


12. A game like Dungeons and Dragons may ask a player to roll two 20-sided dice and take an action based on the maximum of the rolls. Write a for loop that rolls two 20-sided dice 1,000 times and records the maximum of the rolls each time. Then create a graph displaying how often each sum occurs. What do you notice?

```{r}

```

**Answer:**

<br>


13. (Hard Bonus) In a game like Yahtzee you roll five dice. There are many ways to score points, but a Yahtzee is achieved when all five dice are the same number. A player can re-roll any number of dice two times (for three total rolls). (Note: You will not be able to use your current function to do this. You might be able to update it or to use the `sample()` function. Functions like `unique()` and `length()` can help you check if you have duplicates in your roll.)

a. Use conditional computing to simulate rolling dice three times in a way that maximizes your chances of getting five dice of the same number.

```{r}

```


b. Put this code into a loop. Run 1,000 simulations to estimate how often someone would get a Yahtzee using this approach.

```{r}

```

