---
title: 'Activity 7 - Exploring the Central Limit Theorem'
author: "Dr. VanKrevelen -- STS3470 (Spring 2023)"
date: 'Updated: `r Sys.Date()`'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:** Use this .RMD script to record your code / output / answers.

Remember that with a .RMD file, you can run individual code chunks by clicking the green triangle button in the top right of the chunk or by putting your cursor on a line and hitting Cmd+Return (Mac) or Ctrl+Return (PC) like you would in a .R script. When you use the Knit button, it will run all the code in order in a fresh R session and produce your knitted file.

Have questions? Find a classmate or Dr. V to ask!

# Part A - The Central Limit Theorem

1. Generate a matrix of 1000 rows and 10 columns where each cell contains a value from a normal random variable with a mean of 50 and a standard deviation of 4. Call your matrix `mymatrix`. (Note: To check if it worked, see if the mean and sd of your matrix are close to 50 and 4)


```{r}


```

**Mean = **

**SD = **


<br>
<br>


2. I've written code below that will calculate the mean of each row of your matrix. You may want to view the help file for the `apply()` function to see if you can figure out how this code works. After that, add code that creates a histogram of your sample means (called `xbar` in our data frame). (Note: You will need to change `eval = FALSE` to `eval = TRUE` for this code to run when you knit your document.) 

```{r, eval = FALSE}
mymeans <- data.frame(row = 1:1000)
mymeans$xbar <- apply(mymatrix, 1, mean)
```

a) Where is your graph of sample means centered?

**Answer:**

<br>

b) What shape does your distribution of sample means have?

**Answer:**

<br>

c) How does the spread of the sample means differ from the original population you simulated them from?

**Answer:**

<br>
<br>


3. The code below plots our 1,000 sample means along with our original population (solid blue line) and a normal distribution with mean 50 but a standard deviation of only 4 over square root of 10 (dashed red line). Which distribution matches our distribution of *sample means* better? (Again, you will need to change `eval` to be `TRUE` for this to show up in your knitted file.)

```{r, eval = FALSE}
ggplot(mymeans, aes(x = xbar)) + 
  geom_histogram(aes(y = after_stat(density)), color = "white", binwidth = 0.5) + 
  geom_function(fun = dnorm, n = 101, args = list(mean = 50, sd = 4), 
                color = "blue") + 
  geom_function(fun = dnorm, n = 101, args = list(mean = 50, sd = 4 / sqrt(10)), 
                color = "red", linetype = 2)
```

**Answer:**


<br>
<br>


4. Let's try making a QQ Plot to compare our distribution of sample means to a $N(50, \frac{4}{\sqrt{10}})$ distribution. What do you notice? Does our distribution of sample means seem to follow this distribution?

```{r, eval = FALSE}
ggplot(mymeans, aes(sample = xbar)) + 
  geom_qq(distribution = qnorm, dparams = list(mean = 4, sd = 4 / sqrt(10))) +
  geom_qq_line(distribution = qnorm, dparams = list(mean = 4, sd = 4 / sqrt(10))) +
  labs(title = "Comparing sample means to Normal(50, 10/sqrt(4)) dist.",
       subtitle = "Samples of 10 from Normal(50, 10) dist.")
```

**Answer:**


<br>
<br>


**Note**: We just saw that when we take a bunch of random samples from a normal distribution, the distribution of *sample means* is also normal with the same mean but with a smaller standard deviation. Does this work for other distributions?? Let's try it out!

<br>
<br>


5. Use the `rchisq()` function to generate a matrix of 1000 rows and 30 columns where each cell contains a value from a Chi-Squared distribution with between 2 and 10 degrees of freedom (similar to how we created our matrix of values from a normal distribution in question 1). Call your matrix `mymatrix2`. (Note: A Chi-squared distribution has a mean equal to its df and a standard deviation equal to sqrt(2 * df). Use this information to check that you properly generated your values.)

```{r}


```

**Mean = **

**SD = **

<br>
<br>


6. Use the example in question 2 to create a data frame that includes the sample mean for each row of your matrix (each sample of 30 values). Call your data frame `mymeans2`. Create a histogram of those sample means.

```{r}

```

a) Where is your graph of sample means centered? Is it close to the degrees of freedom?

**Answer:**

<br>

b) What shape does your distribution of sample means have? Is it more similar to your population (Chi-Squared) or a Normal distribution?

**Answer:**

<br>

c) Is the standard deviation of your sample means closer to the standard deviation of the population (sqrt(2 * df)) or to sqrt(2 * df) / sqrt(30)? Why do you think this is?

**Answer:**

<br>
<br>


7. Change the first line of the code below to assign df the value you used for the degrees of freedom in your Chi-Squared distribution. Then run the code (changing `eval` to `TRUE`). How well does the normal distribution curve match your graph of sample means?

```{r, eval = FALSE}
df <- 5
ggplot(mymeans2, aes(x = xbar)) + 
  geom_histogram(aes(y = after_stat(density)), color = "white", bins = 15) + 
  geom_function(fun = dnorm, n = 101, color = "red",
                args = list(mean = df, sd = sqrt(2 * df / 30)))
```

**Answer:**

<br>
<br>


8. Lastly, let's use a QQ plot to compare our sample means to a normal distribution.

```{r, eval = FALSE}
ggplot(mymeans2, aes(sample = xbar)) + 
  geom_qq(distribution = qnorm) +
  geom_qq_line(distribution = qnorm) +
  labs(title = "Are sample means from Chi-Squared distributions normally distributed?",
       subtitle = paste0("Samples of 30 from a Chi-Squared(", df, ") distribution"))
```


**Note:** We have seen two cases where the distribution of sample means followed a normal distribution. In one case our population was normally distributed, and in the other case it was not.

---

# Part B - Practice finding p-values / areas under distributions

This part contains several problems for you to practice using functions related to distributions in R. For each problem, you are strongly encouraged to sketch a picture to verify that your answer makes sense. You could also use a tool like StatKey or SAS to verify your answers.

**Notation**: In statistics, the notation P(Z > 1.28) means "probability that a random value from a Z distribution is above 1.28". In other words, you are being asked to find the area above 1.28 on a Z distribution. You may see Z replaced by other distributions like T, F, $\chi^2$, or other normal distributions.

1) Write R code to find P(Z > 1.28) and fill in the answer below.

```{r}

```

**P(Z > 1.28) =**

<br>


2) Write R code to find P(Z > -1.28) and fill in the answer below.

```{r}

```

**P(Z > -1.28) =**

<br>


3) Earlier in Notes 07, we found $P(T_{999} > 1.28) = 0.1004$. How does this compare to your answer from question 1 and what does this demonstrate about the relationship between the Z and T distributions?

**Answer:**

<br>


4) Randomly sample 1000 values from a $T$ distribution with 999 d.f. and use a QQ plot to compare them to a standard normal distribution (see code from earlier in Notes 7). How does this graph help support your answer to the previous question?

```{r}

```

**Answer:**

<br>


5) Write R code to find P(X > 1.8) where X has a normal distribution with mean 1.5 and standard deviation 0.3. Fill in the answer below.

```{r}

```

**P(X > 1.8) =**

<br>


6) Write R code to find P(0.9 < X < 1.8) where X has a normal distribution with mean 1.5 and standard deviation 0.3. Fill in the answer below. (Hint: You will need to use a function more than once and put your results together.)

```{r}

```

**P(0.9 < X < 1.8) =**

<br>


7) Write R code to find $P(F_{12, 6} < 1.3)$ and fill in the answer below.

```{r}

```

**$P(F_{12, 6} < 1.3)$ =**

<br>


8) Write R code to find $P(F_{12, 6} > 1.3)$ and fill in the answer below.

```{r}

```

**$P(F_{12, 6} > 1.3)$ =**

<br>


9) Write R code to find the value that we should replace the ? with to make this a true statement: $P(T_{12} < ?) = 0.975. Report your answer below.

```{r}

```

**$P(T_{12} < ?) = 0.975$**

<br>


10) Write R code to find the value that we should replace the ? with to make this a true statement: $P(Z > ?) = 0.05. Report your answer below.

```{r}

```

**$P(Z > ?) = 0.05$**

<br>