---
title: "Fancy tables with kable and kableExtra"
author: 'STS3470: Dr. VanKrevelen'
date: "Tuesday, April 4, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In many cases you may want to display data in a table for a reader, but the default options aren't great. For starters, the output is displayed with ## at the start of each line.

```{r}
women
```

The `kableExtra` package can help!

```{r}
library(kableExtra)
kbl(women)
```

The default option still doesn't look great... However, there are a series of functions called `kable_xxxx()` where xxxx can be styling, paper, classic, classic2, minimal, etc. Below are a few examples. Notice that you can also add captions in the `kbl()` function.

```{r}
kbl(women, caption = "A table using styling") %>%
  kable_styling()

kbl(women, caption = "This one has a paper theme") %>%
  kable_paper()

kbl(women, caption = "And a classic theme") %>%
  kable_classic()
```

You may find that it looks weird with so much blank space. We can use the `full_width = FALSE` argument with any of these functions. We can also add `bootstrap_options` argument with a "striped" (shade every other cell) and/or hover (shade cells your cursor is over)  option.

```{r}
kbl(women) %>%
  kable_classic("hover", full_width = FALSE)
```

We can choose where to position the table and make changes to things like font size. We can add `column_spec()` functions to change things like column colors.

```{r}
kbl(women) %>%
  kable_classic("striped", full_width = FALSE, font_size = 16) %>%
  column_spec(column = 1, background = "violet", border_left = TRUE, 
              italic = TRUE) %>%
  column_spec(column = 2, color = "blue", border_right = TRUE, bold = TRUE)
```

If you are knitting to html, you can also make the table scrollable to save space. I've illustrated this with a larger data frame below to show you what it looks like. The height argument uses number of pixels and may take some trial and error.

```{r}
kbl(USJudgeRatings) %>%
  kable_paper("striped", full_width = TRUE) %>%
  scroll_box(width = "75%", height = "400px")
```

For more examples including MANY more things you can do, check out [this guide](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html).